---
title: "ESM 244 Week 5 Lab"
author: "Meghan Fletcher"
date: "2/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tsibble)
library(feasts)
library(fable)
library(here)
```

## Read in energy data and convert to a tsibble

```{r}
energy <- read_csv(here("data", "energy.csv"))
```

1. Add a new column (date) that is the current month column converted to a time series class, yearmonth
2. Convert the data frame to a tsibble, with that date column as the time index
```{r}
energy_ts <- energy %>% 
  mutate(date = tsibble::yearmonth(month)) %>% 
  as_tsibble(key = NULL, index = date)
```

## Exploratory time series visualization

```{r}
ggplot(data = energy_ts, aes(x = date, y = res_total)) +
  geom_line() +
  labs(y = "Residentail energy consumption \n (Trillion BTU)")
```
**Take-aways from this plot:**
 - Overall increasing trend overall, but stability (and possibly a slight decreasing trend) starting around 2005
 - Clear seasonality, with a dominant seasonal feature and also a secondary peak each year - that secondary peak has increased substantially
 - No notable cyclicality or outliers
 
### Seasonplot:
```{r}
# use feasts::gg_season() to create an exploratory seasonplot, which has month on the x-axis, energy consumption on the y-axis, and each year is its own series (mapped by line color).

energy_ts %>% 
  gg_season(y = res_total) +
  theme_minimal() +
  labs(x = "month",
       y = "residential energy consumption (trillion BTU)")
```

**Take-aways from this plot:**
 - The highest residential energy usage is around December / January / February
 - There is a secondary peak around July & August (that’s the repeated secondary peak we see in the original time series graph)
 - We can also see that the prevalence of that second peak has been increasing over the course of the time series: in 1973 (orange) there was hardly any summer peak. In more recent years (blue/magenta) that peak is much more prominent.

### Subseries plot:
```{r}
energy_ts %>% gg_subseries(res_total)
```
**Take-away here:** there is clear seasonality (higher values in winter months), with an increasingly evident second peak in June/July/August. This reinforces our takeaways from the raw data and seasonplots.


### Decomposition (here by STL)

STL decomposition allows seasonality to vary over time (a major difference from classical decomposition, and important here since we do see changes in seasonality).

```{r}
# Find STL decomposition
dcmp <- energy_ts %>% 
  model(STL(res_total ~ season()))

# View the ocmponents: components(dcmp)

# Visualize the decomposed components
components(dcmp) %>%  autoplot() +
  theme_minimal()
```

### Autocorrelaiton function (ACF)
We use the ACF to explore autocorrelation (here, we would expect seasonality to be clear from the ACF):

```{r}
energy_ts %>% 
  ACF(res_total) %>% 
  autoplot()
```

We see that observations separated by 12 months are the most highly correlated, reflecting strong seasonality we see in all of our other exploratory visualizations.


## Forecasting by Holt_Winters exponential smoothing

To create the model below, we specify the model type (exponential smoothing, ETS), then tell it what type of seasonality it should assume using the season("") expression, where “N” = non-seasonal (try changing it to this to see how unimpressive the forecast becomes!), “A” = additive, “M” = multiplicative. Here, we’ll say seasonality is multiplicative due to the change in variance over time and also within the secondary summer peak:

```{r}
# Create the model
energy_fit <- energy_ts %>% 
  model(
    ets = ETS(res_total ~ season("M"))
  )

# Forecast using the model 10 years into the future:
energy_forecast <- energy_fit %>% 
  forecast(h = "10 years")

#Plot just the forecasted values (with 80 & 95% CIs):
energy_forecast %>% 
  autoplot()
```

```{r}
# Or plot it added to the original data:
energy_forecast %>% 
  autoplot(energy_ts)
```

### Assessing residuals:








